---
title: "PracticalMachineLearning_ROBMILL"
output: html_document
---
### Final Peer Reviewed Assignment
## Practical Machine Learning
## Week 4

```{r preparation, results='hide'}
# Load Librarys
# Load Data
library(kernlab)
library(caret)
library(ggplot2)
library(Hmisc)
library(gridExtra)
library(graphics)

training<-read.csv(file = "pml-training.csv",stringsAsFactors = FALSE) # Read training data

testing<-read.csv(file = "pml-testing.csv",stringsAsFactors = FALSE) # Read testing data
```

### Initial EDA on Raw Data Set
Data comes from: 
1.  <http://groupware.les.inf.puc-rio.br/har>
2.  <http://groupware.les.inf.puc-rio.br/har#weight_lifting_exercises>

The data set describes quality of execution for the wearer while doing dumbell lift exercises.  

Different Classes of execution quality.  High level Class A represents an example of successful execution of the exercise while Classes B-E represent examples of common mistakes made.  

* Class A: exactly according to specification.
* Class B: throwing the elbows to the front.
* Class C: lifting the dumbell only halfway.
* Class D: lowering dumbell only halfway.
* Class E: throwing the hips to the front.

### The goal of the exercise is to identify predictors that accurately determine execution quality or manner of exercise.

Similar to the iris data set exercise predicting species based on a set of features.  q
```{r EDA}

dim(training) # display number or rows/columns
# str(training) # display features
# summary(training) # display summary stats
#names(training)
head(training)
table(training$classe)

qplot(training$total_accel_forearm,training$total_accel_belt,colour=classe,data=training)

qplot(training$total_accel_arm,training$total_accel_dumbbell,colour=classe,data=training)

qplot(training$total_accel_arm,training$total_accel_belt,colour=classe,data=training)

qplot(training$total_accel_forearm,training$total_accel_dumbbell,colour=classe,data=training)


featurePlot(x=training[,c("total_accel_forearm","total_accel_dumbbell","total_accel_belt","total_accel_arm")], y=training$classe, plot="pairs")
```


### Partition Training Data and Processing
```{r Data Partitioning}

inTrain<-createDataPartition(training$classe,p=0.70,list=FALSE)

trainingSub<-training[inTrain,]
testingSub<-training[-inTrain,]

modFit<-train(classe~total_accel_dumbbell+total_accel_arm,method="rpart",data=training)
print(modFit$finalModel)
```

## Citation
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: <http://groupware.les.inf.puc-rio.br/har#weight_lifting_exercises#ixzz4aHsvXnQK>