---
title: "PracticalMachineLearning_ROBMILL"
output: html_document
---
### Final Peer Reviewed Assignment
## Practical Machine Learning
## Week 4

```{r preparation}
# Load Librarys
# Load Data
library(kernlab)
library(caret)
library(ggplot2)
library(Hmisc)
library(gridExtra)

training<-read.csv(file = "pml-training.csv") # Read training data

testing<-read.csv(file = "pml-testing.csv") # Read testing data
```

### Initial EDA on Raw Data Set
Data comes from <http://groupware.les.inf.puc-rio.br/har>
<http://groupware.les.inf.puc-rio.br/har#weight_lifting_exercises>

### The data set seeks to identify quality of execution for the wearer while doing dumbell lift exercises.  

### Different Classes of execution quality
* Class A: exactly according to specification.
* Class B: throwing the elbows to the front.
* Class C: lifting the dumbell only halfway.
* Class D: lowering dumbell only halfway.
* Class E: throwing the hips to the front.
```{r EDA}

dim(training) # display number or rows/columns
str(training) # display features
summary(training) # display summary stats
names(training)

plot(training$classe,type="hist")
```

### Partition Training Data and Processing
```{r }

inTrain<-createDataPartition(training$classe,p=0.70,list=FALSE)

trainingSub<-training[inTrain,]
testingSub<-training[-inTrain,]


```